Models of gravitational lens candidates from Space Warps CFHTLS

### 2017-06-29 paper2.tex
adding comments about SW57, fixes #25
forgotten last minute changes on main.tex
continuing response

%%% DEBUG SETTINGS - remove at the end %%%%%%%%%%%%%%%
\usepackage{silence} % silence error warinings
\WarningFilter{caption}{Unsupported document class} % caption doesn't know mnras..
\hbadness = 10000
\vbadness = 10000
% \usepackage{}

@@

\begin{abstract}
We report modelling follow-up of recently-discovered
gravitational-lens candidates in the Canada France Hawaii Telescope
Legacy Survey. Lens modelling was done by a small group of
specially-interested volunteers from the \SW citizen-science community
who originally found the candidate lenses.  Models are categorised
according to seven diagnostics indicating (a)~the image morphology and
how clear or indistinct it is, (b)~whether the mass map and synthetic
lensed image appear to be plausible, and (c)~how the lens-model mass
compares with compares with the abundance-matched halo mass derived
from photometry of the lensing galaxy.

@@

\begin{keywords}
gravitational lensing: strong -- citizen science
\end{keywords}

>>>>

%%% DEBUG SETTINGS - remove at the end %%%%%%%%%%%%%%%

%\usepackage{silence} % silence error warinings
%\WarningFilter{caption}{Unsupported document class} % caption doesn't know mnras..
%\hbadness = 10000
%\vbadness = 10000

% \usepackage{}

@@

\begin{abstract}
We report modelling follow-up of recently-discovered
gravitational-lens candidates in the Canada France Hawaii Telescope
Legacy Survey. Lens modelling was done by a small group of
specially-interested volunteers from the \SW\ citizen-science community
who originally found the candidate lenses.  Models are categorised
according to seven diagnostics indicating (a)~the image morphology and
how clear or indistinct it is, (b)~whether the mass map and synthetic
lensed image appear to be plausible, and (c)~how the lens-model mass
compares with compares with the abundance-matched halo mass derived
from photometry of the lensing galaxy.

@@

\begin{keywords}
gravitational lensing: strong -- galaxies: general
\end{keywords}

### 2017-06-29 parts/app.tex
adding comments about SW57, fixes #25
liewrap
continuing response

The models of simulated lenses in \cite{2015MNRAS.447.2170K} showed a
tendency to produce too shallow density profiles.
Allowing smaller mass tiles in the
central region, thus allowing the mass profile to rise more steeply
near the centre, was suggested as a possible cure.

@@

\subsection{Parameterisation of pixel models} \label{subsec:parameter}

In order to fit the set of pixelated models to a single parameterised model, a program was written that took a parameterised function and subtracted from it the mean and the principal components of the data, which were calculated using classical Principal Component Analysis.
This created the residuals function.
The number of components used in the analysis was varied, to test how this affected the output, and it was found that using 5 principal components tended to give a reasonable approximation.
A masking function was added which selected only the data points that fell inside the image of the lens, and the principal components were clipped in order to keep the values inside the region of the ensemble of models.
Any value higher than the clip was set to be the clip value.
This was chosen to be 2.5 as, assuming that the data follows a Gaussian error distribution, almost all the values for the variance should lie between 2 and 3 standard deviations from the mean.
Minimising the residuals function produces the set of parameters that fit the parameterised function to the original pixelated ensemble most closely.
A least squares fit was used to perform this minimisation.
The parameterised model function was obtained from the gravitational potential of an isothermal ellipsoid mass distribution \citep{2001astro.ph..2341K}.
This model is frequently used to describe gravitational lenses as it tends to fit well with observations.
The isothermal ellipsoid model outputs three useful parameters: the radius of the Einstein ring, the ellipticity of the model and the angle of the ellipticity from the vertical, giving the orientation of the galaxy.
By applying this model to simulated lenses for which the values of these parameters were already known, it was possible to gain an estimate of the projected accuracy of the results, before applying the model to the candidate lensing galaxies.

@@

\begin{figure}
  \includegraphics[width=\linewidth]{img/rE_comp/rE_comp.png}
  \caption{
    Comparison of Einstein radii \ER obtained from mass tiles directly to those obtained from a parameterized model to test the performance of the parametrisation algorythm.
    The parameterized model was generated using princible componant analysis on the ensemble of models.
    The blue dashed line represents a perfect recovery of \ER.
    }
  \label{fig:parameter}
\end{figure}

>>>>

The models of simulated lenses in \cite{2015MNRAS.447.2170K} showed a tendency 
to produce density profiles which were too shallow, resulting typically
in overestimates of the Einstein radius.
Allowing smaller mass tiles in the
central region, thus allowing the mass profile to rise more steeply
near the centre, was suggested as a possible cure.

@@

\subsection{Parametrization of pixel models} \label{subsec:parameter}

In order to fit the set of pixelated models to a single parametrized model, a 
program was written that took a parametrized function and subtracted from it 
the mean and the principal components of the data, which were calculated using 
classical Principal Component Analysis.
This created the residuals function.
The number of components used in the analysis was varied, to test how this 
affected the output, and it was found that using 5 principal components tended 
to give a reasonable approximation.
A masking function was added which selected only the data points that fell 
inside the image of the lens, and the principal components were clipped in 
order to keep the values inside the region of the ensemble of models.
Any value higher than the clip was set to be the clip value.
This was chosen to be 2.5 because, assuming that the data follows a Gaussian error 
distribution, almost all the values for the variance should lie between 2 and 3 
standard deviations from the mean.
Minimising the residuals function produces the set of parameters that fit the 
parametrized function to the original pixelated ensemble most closely.
A least squares fit was used to perform this minimisation.
The parametrized model function was obtained from the gravitational potential 
of an isothermal ellipsoid mass distribution \citep{2001astro.ph..2341K}.
This model is frequently used to describe gravitational lenses as it tends to 
fit well with observations.
The isothermal ellipsoid model outputs three useful parameters: the radius of 
the Einstein ring, the ellipticity of the model and the angle of the 
ellipticity from the vertical, giving the orientation of the galaxy.
By applying this model to simulated lenses for which the values of these 
parameters were already known, it was possible to gain an estimate of the 
projected accuracy of the results, before applying the model to the candidate 
lensing galaxies.

@@

\begin{figure}
  \includegraphics[width=\linewidth]{img/rE_comp/rE_comp.png}
  \caption{
    Comparison of Einstein radii \ER\ obtained from mass tiles directly to 
    those obtained from a parameterized model to test the performance of the 
    parametrization algorithm.  The parameterized model was generated using 
    principal component analysis on the ensemble of models.  The blue dashed
    line represents a perfect recovery of \ER.
    }
  \label{fig:parameter}
\end{figure}
